Reproducible, flexible and high throughput data extraction from primary literature: The \pkg{metaDigitise} \proglang{R} package

### 2017-11-29 docs/manuscript.tex
Corrections to MS based on Shin's comments, some really good feedback

%% - \Abstract{} almost as usual
\Abstract{

}

@@

In many different contexts, researchers need to make use of data presented in primary literature. Most notably, this includes meta-analysis, which is becoming increasingly common in many research fields. Meta-analysis uses effect size estimates and their sampling variance, taken from many studies, to understand whether particular effects are common across studies and to explain variation among these effects \citep{Glass1976,Borenstein2009,Koricheva2013,Naka2017}. Meta-analysis therefore relies foremost on data extracted from primary literature and more specifically statistics that have been reported in the text or tables of research papers. These summary statistics are also, however, frequently presented in figures and so need to be manually extracted using digitising programs. Although there are several existing tools to perform tasks like this (e.g. \proglang{DataThief} \citep{DataThief}, \proglang{GraphClick} \citep{GraphClick}, \proglang{WebPlotDigitizer} \citep{WebPlotDigitizer}), these tools are not designed specifically for meta-analysis. 
%% need to spell out clear points
%% 1 - dont manipulate data/allow different plot types

@@

When \fct{metaDigitise} is run, it recognizes all the images in a directory and automatically imports them one by one, allowing the user to click and enter relevant information about a figure as they go. This expedites digitising figures by preventing users from having to constantly change directories and / or open new images. The data from a completed image is automatically saved as a \code{metaDigitise} object in an .RDS file to a \code{caldat} directory, that is created within the parent directory when first executing the \fct{metaDigitise} function. These files enable re-plotting and editing of images at a later point (see below).

@@

Although \pkg{metaDigitise} is already very flexible, and provides functionality not seen in any other packages (Table \ref{tab:comparison}) it is clear that there are some functions that it does not perform. A notable feature that \pkg{metaDigitise} lacks is automated point detection. This is available in several packages (Table \ref{tab:comparison}). However, from our experience of using these functions, manual digitising is more reliable and often equally as fast. Particularly given that calibration (for point detection) needs to be done for each plot individually in any case. Additionally, auto-detection often misses many points which then subsequently need to be manually added. Based on tests of \pkg{metaDigitise} (see above) figures can be extracted in around 1-2 minutes, including the entry of metadata. As a result, we do not belive that current automated point detection provides substantial benefits in terms of time or accuracy.

>>>>

%% - \Abstract{} almost as usual
\Abstract{
Research synthesis, especially, in the form of meta-analysis requires data extraction from primary studies. Meta-analysis synthesizes effect sizes, often calculated from summary statistics of studies. However, exact values of such statistics are commonly hidden in figures. The R package metaDigitise extracts descriptive statistics such as means, standard deviations and, if applicable, correlations from the four types of plots: 1) mean and error plots (e.g. bar graphs with standard errors), 2) box plots, 3) scatter plots and 4) histograms. The package interactively guides the user through data extraction process. Notably, it enables a large-scale extraction using image files, letting the user stop processing, and edit and add to the resulting data fame at any point. Further, it facilitates reproducible data extractions from plots with little inter-observer bias, thus, allowing a group of people to participate the extraction of data collaboratively.
}

@@

In many different contexts, researchers need to make use of data presented in primary literature. Most notably, this includes meta-analysis, which is becoming increasingly common in many research fields. Meta-analysis uses effect size estimates and their sampling variance, taken from many studies, to understand whether particular effects are common across studies and to explain variation among these effects \citep{Glass1976,Borenstein2009,Koricheva2013,Naka2017}. Meta-analysis therefore relies foremost on data extracted from primary literature, and more specifically, statistics that have been reported in the text or tables of research papers (e.g., means, standard deviations, correlation coefficients). These summary statistics are also, however, frequently presented in figures and so need to be manually extracted using digitising programs. While inferential statistics (e.g., t- and F-statistics) are often presented along side summary statistics and used to derive effect sizes, raw descriptive statistics are much more appropriate to use because sources of non-independence in experimental designs can be dealt with more easily using summary statistics \citep{Noble2017}. Although there are several existing tools to perform tasks like this (e.g. \proglang{DataThief} \citep{DataThief}, \proglang{GraphClick} \citep{GraphClick}, \proglang{WebPlotDigitizer} \citep{WebPlotDigitizer}), these tools are not designed specifically for meta-analysis. 
%% need to spell out clear points
%% 1 - dont manipulate data/allow different plot types

@@

When \fct{metaDigitise} is run, it recognizes all the images in a directory and automatically imports them one by one, allowing the user to click and enter relevant information about a figure as they go. This expedites digitising figures by preventing users from having to constantly change directories and / or open new images. The data from a completed image is automatically saved as a \code{metaDigitise} object in an .RDS file to a \code{caldat} directory that is created within the parent directory when first executing the \fct{metaDigitise} function. These files enable re-plotting and editing of images at a later point (see below).

@@

Although \pkg{metaDigitise} is already very flexible, and provides functionality not seen in any other packages (Table \ref{tab:comparison}) it is clear that there are some functions that it does not perform. A notable feature that \pkg{metaDigitise} lacks is automated point detection. Point detection is available in several packages (Table \ref{tab:comparison}). However, from our experience of using these functions, manual digitising is more reliable and often equally as fast. Particularly given that calibration (for point detection) needs to be done for each plot individually in any case. Additionally, auto-detection often misses many points which then subsequently need to be manually added. Based on tests of \pkg{metaDigitise} (see above) figures can be extracted in around 1-2 minutes, including the entry of metadata. As a result, we do not belive that current automated point detection provides substantial benefits in terms of time or accuracy.
